Waving and swaying

### 2013-12-17 Paper/grass.tex
commit2

existence of an inflection point in mean velocity near top of canopy. In addition pattern of coherent eddies propagating downstream 
along have also been observed.

@@

Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 

@@

similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami.
Several aspect of existing theory remain unclear. The proposed hydrodynamic instability of shear layer layer does not take into account presence of grass whereas shear layer 
was produced due to drag caused by presence vegetation. Since shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
atmospheric flow is unbounded in vertical direction.

@@

In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
prediction from theory with existing data

\section{The model}
We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
% and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
\begin{equation}
 \mathbf{f}=-N\mathbf{f_{d}}
\end{equation}
where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, orientation and diameter of grass which we model as 
\begin{equation}
 \mathbf{f_{d}}=C_{N}\rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
\end{equation}

@@

%\end{equation}
where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.

@@

In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow, 
we investigate stability of base flow profile which can be obtained by solving steady state solution of equation(1). We consider small perturbations $u, v, p$  associated 
with base profile $U$ and $P$ respectively. Momentum and mass balance equation expanded in 1st order of perturbed variable yields.

@@

We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{c_{1}H^2}{4\mu} \]
where $c_{1}$ is constant pressure gradient, H is height at which fluid is present. With these scaling and using stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations 
can be simplified into a single equation.

>>>>

presence of an inflection point in mean velocity near canopy top.
% In addition pattern of coherent eddies propagating downstream 
%along have also been observed.

@@

%\newline
Following the work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 

@@

similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami. While shear layer
model sucessfully predict frequency of passage of corent vortices for number of experimental observations,
several aspect of existing theory remain unclear. First, the proposed hydrodynamic instability of shear layer does not take into account influence of
drag due to grasses whreas shear layer itself is produced by drag force provided by vegetation. 
%the proposed hydrodynamic instability mechanism of shear layer layer does not take into account presence of grass whereas existence of 
%shear layer is due to drag provided by vegetation. 
Second, classical free shear flow is known to be unstable above a very small Reynolds number $\geq 5 $. On the contrary experimental observations
(Grizzle,1996) shows that an event of mo-nami is not observed untill flow speed exceed a certain threshold value. Although atmoshpheric flow through terrestrial canopy seems similar to 
aquatic flow through vegetation, atmoshpheric flow is essentially unbounded, due to difference of vertical extent of flow in terrestrial and aquatic case a mechanism similar to 
terrestrial case might not be adequate for aquatic scenario.
%might not hold  
%was produced due to drag caused by presence vegetation. 
%Second, shear layer instability mechanism does not predict any threshold for monami, hence it fails to explain
%observations where monami is observed when mean flow speed is above certain value. While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, 
%atmospheric flow is unbounded in vertical direction.

@@

 In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegeation. Our caluation of linear stability analysis of flow in presence of grasses 
shows that drag from the grass blades damps out the Kelvin Helmholtz instability, normally observed in a free shear layer. We also show that an event of mo-nami is caused by instablity 
of fluid flow above vegeation rather than that of shear layer near grass. We further show that inclusion of grass predicts a threshold criteria for waving characterizedd by
Reynolds number. We confirm our analyis by comparing predictin from theory with field and experimental observations.  
%In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
%medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
%a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
%We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
%prediction from theory with existing data

\section{The model}
We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continous drag filed which is taken as additional body force 
in momentum equation of fluid. 
%We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
%taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
%shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
% and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
%\begin{equation}
% \mathbf{f}=-N\mathbf{f_{d}}
%\end{equation}
where $\mathbf{f}=-N\mathbf{f_d}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
%\begin{equation}
% \mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
%\end{equation}\

@@

%\end{equation}
%where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.

@@

In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of hydrodynamic instability associated with mean flow, 
we investigate stability of base flow profile which obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 
with base profile $U$ and $P$ respectively, momentum and mass balance equation expanded in 1st order of perturbed variable yields.

@@

%We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 
%flow with zero shear at top. We use following scaling to non-dimensionlize equation of motion
Where drag coeffient $C_N$ is set zero out side vagetation. Using height $H$ of fluid column and associated advection time H/$U_{0}$, we use following scalaing to non-dimensinlize equation 
of motion
   \[ u = U_{0}\bar{u},\hspace{1cm} y = (H/2)\bar{y}, \hspace{1cm} \text{with}\hspace{2mm} U_{0}=\frac{(dP/dx)H^2}{4\mu} \]
With these scaling along with use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.
